PhD Title \\ PhD Thesis

### 2024-06-10 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:

@@

These results are somewhat contrary to other work with silica dimers
\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
have trapped the dimer in an orientation perpendicular to the beam
propagation direction.  The rotational motion is induced due to the
asymmetric geometry creating an unbalanced polarisation susceptibility
along its long axis as compared to its short axis; therefore its long
axis is aligned with the polarisation vector and can rotate
freely\cite{Ahn2018}.  This however cannot be the case with our
simulations as the dimer rotates about its long axis, meaning there
cannot be an asymmetric axis to align with the beam's polarisation
vector. Furthermore, we see a non-linear increase in the rotational
speed of our dimers with size, the drag torque from the surrounding
fluid is $\propto r^3$ so the expectation is that the rotation
frequency should fall off with increasing size.  This indicates that
the rotational motion is due to the shape asymmetry of the dimer and
not solely due to the beam's angular momentum.  Measurement of this
photo-kinetic force is difficult to achieve due to the fact that
previous analysis was conducted in the Rayleigh regime, where the
polarizability of our dimer can be approximated as:
\begin{align}
  \bold{p}({\bf r},t)
  =
  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
\end{align}
where the polarizability is given as a 3D vector for the three
principle Cartesian directions.  In order to measure the magnitude of
second order contributions we would need to construct a dipole array
that fully captures the scattering of a dimer. Measuring the optical torque 
makes it clear that the polarizability is a contributing factor to this optical 
rotation phenomena. Rotating a symmetric dimer in the $x-z$ plane reveals 
that while the dimer can be rotated in an orientation perpendicular 
to the beam rotational torque is maximised when rotated while aligned with 
the optical axis.

\begin{figure}[t]
	\begin{subfigure}{\linewidth}
	\centering
	\includegraphics[width=0.6\linewidth]{tx_different_polarisations.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
	\includegraphics[width=0.6\linewidth]{ty_different_polarisations.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
	\includegraphics[width=0.6\linewidth]{tz_different_polarisations.png}
	\end{subfigure}
\caption{Calculation of the optical torque about the three primary axis on a symmetric dimer in linear, elliptical, and circular polarisation beams.}
\end{figure}
\newpage

\comment{It would be good to have a plot of the torque in each of the
  three axes as a function of the polarization of the trapping beam,
  from linearly polarized to elliptically polarized to circularly polarized.}

>>>>

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}
\newpage
\subsection{Non-trivial harmonic traps}\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:

@@

These results are somewhat contrary to other work with silica dimers
\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
have trapped the dimer in an orientation perpendicular to the beam
propagation direction.  The rotational motion is induced due to the
asymmetric geometry creating an unbalanced polarisation susceptibility
along its long axis as compared to its short axis; therefore its long
axis is aligned with the polarisation vector and can rotate
freely\cite{Ahn2018}.  This however cannot be the case with our
simulations as the dimer rotates about its long axis, meaning there
cannot be an asymmetric axis to align with the beam's polarisation
vector. Furthermore, we see a non-linear increase in the rotational
speed of our dimers with size, the drag torque from the surrounding
fluid is $\propto r^3$ so the expectation is that the rotation
frequency should fall off with increasing size.  This indicates that
the rotational motion is due to the shape asymmetry of the dimer and
not solely due to the beam's angular momentum.  Measurement of this
photo-kinetic force is difficult to achieve due to the fact that
previous analysis was conducted in the Rayleigh regime, where the
polarizability of our dimer can be approximated as:
\begin{align}
  \bold{p}({\bf r},t)
  =
  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
\end{align}
where the polarizability is given as a 3D vector for the three
principle Cartesian directions.  In order to measure the magnitude of
second order contributions we would need to construct a dipole array
that fully captures the scattering of a dimer. Measuring the optical torque 
makes it clear that the polarizability is a contributing factor to this optical 
rotation phenomena. Rotating a symmetric dimer in the $x-z$ plane reveals 
that while the dimer can be rotated in an orientation perpendicular 
to the beam rotational torque is maximised when rotated while aligned with 
the optical axis.
\begin{SCfigure}
	\includegraphics[width=0.7\linewidth,height= 22cm]{torque_different_polarisations.png}
	\caption{Optical torque against polar angle $\theta$ about the three primary axis (top: torque about the x-axis; middle: torque about the y-axis; bottom: torque about the z-axis) on a symmetric dimer in linear, elliptical, and circular polarisation beams. Diagram to the right is for visual clarity about the direction of $\theta$. \vspace{7.5cm}}
\end{SCfigure}
\newpage

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there is the potential for non-vertical trapping orientations in which the dimer is located in a harmonic trap. When trapped by a circularly polarised light these dimers exhibit gyroscopic precession, demonstrating rotational motion not just about its long axis, but also around the optical axis of the beam. This gyroscopic motion has been demonstrated previously in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed for micron scale aggregates, typically because it was assumed that an off-axis orientation would not be stable due to the magnitude of the electric field.

\comment{Need to add figure here demonstrating the rotational behaviour but not sure how to best convey that motion. Perhaps just a simple trajectory diagram showing the particles motion (position in the first column, $u_x$ in the second, then $u_y$, then $u_z$).}
